---
title: "ABC-XYZ Premise and Dataset Creation"
author: Jonah Lyon (jonah@a2-ai.com)
date: today
date-format: long
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

The purpose of this script is to create a dataset with an imaginary patient population for a future study.

```{r, warning = FALSE, echo = FALSE, message = FALSE}
renv::snapshot()
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(mrgsolve)
library(yspec)

# data spec yaml file
spec <- ys_load("../data/derived/ppk.yaml")
derDir <- "../data/derived/dose_histories"


```

# Load PopPK Model Run for ABC-XYZ
```{r}
mod1 <- mread("../model/R-mrgsolve/1001.mod", 
              start = 0, end = 120, delta = 1)

```

# Create population
- Study of N = 120
- pancreatic cancer (HV = 0)
- ECOG of 0 (45%) and 1 (55%)
- obese study BMI > 30
- hepatic dysfunciton (mild n = 40, moderate n = 60)
- U.S Study
      - 80% White, 15% Black, 3% Asian, 2% Other
      - 52% Female, 48% Male
- QOD, 4.8mg (n = 80) and 7.2mg (n = 40) Dosing Groups.  All doses fasted.

# Define Cohorts and Subjects
```{r}
drug <- "ABC-XYZ"
site <- "001"
cohort1 <- str_c("01", str_pad(1:80, pad = 0, width = 3)) # 4 mg
cohort2 <- str_c("02", str_pad(1:40, pad = 0, width = 3)) # 8 mg

cohort1_usubjid <- str_c(drug, site, cohort1, sep = "-")
cohort2_usubjid <- str_c(drug, site, cohort2, sep = "-")

```
# Random Simulation for subject covariates
## Demographics
```{r}
set.seed(20240107)
subject_dm <- data.frame(
  USUBJID = c(cohort1_usubjid, cohort2_usubjid),
  COHORT = c(rep(1, 80), rep(2, 40)),
  SEXC = sample(c("F", "M"), size = 120, 
                replace = TRUE, prob = c(0.52, 0.48)),
  BBMI = round(rnorm(n = 120, mean = 34.2, sd = 1.4), 1),
  RACEC = sample(c("White", "Black", "Asian", "Other"), size = 120, 
                replace = T, prob = c(0.8, 0.15, 0.03, 0.02)),
  BAGE = round(rnorm(n = 120, 62, sd = 4.7)),
  BECOG = sample(c(0, 1), size = 120, replace = TRUE, prob = c(0.45, 0.55))
) |> 
  mutate(
    SEXF = ifelse(SEXC == "M", 0, 1),
    RACEN = case_when(
      RACEC == "White" ~ 1,
      RACEC == "Black" ~ 2,
      RACEC == "Asian" ~ 3,
      RACEC == "Other" ~ 4
    ),
    BHT = ifelse(SEXF == 1,
                 round(rnorm(n = 120, mean = 163, sd = 6.35)/100, 2),
                 round(rnorm(n = 120, mean = 175, sd = 7.62)/100, 2)),
    BWT = round(BHT^2*BBMI, 1),
    PTYPEC = "Pancreatic Cancer",
    PTYPE = 4
  )

```

## Labs
```{r}
subject_lb <- data.frame(
  USUBJID = c(cohort1_usubjid, cohort2_usubjid),
  BHFCC = sample(c("Mild", "Moderate"), size = 120, replace = TRUE, prob = c(0.4, 0.6)),
  BALB = round(rnorm(n = 120, mean = 3.6, sd = 0.25), 2),
  BCRCL = round(rnorm(n = 120, mean = 85, sd = 12), 1)
) |> 
  mutate(
    BHFC = ifelse(BHFCC == "Mild", 1, 2)
  )

```

## Combine Demographics and Labs
```{r}

subject_dm_lb <- subject_dm |> 
  left_join(subject_lb, by = "USUBJID")

```
# Dosing History
# 28-day cycles
# QOD for 4.8, QOD for 7.2
```{r}
subject_ex <- subject_dm_lb |> 
  select(USUBJID, COHORT) |> 
  mutate(
    AMT = ifelse(COHORT == 1, 4.8, 7.2)
  ) |> 
  crossing(
    ATFD = seq(0, 2016, by = 48) # QOD for 3 28 day cycles, in hours
  ) |> 
  mutate(
    NTFD = ATFD,
    ATLD = 0,
    NTLD = 0,
    EVID = 1, 
    CMT = 1,
    BLQ = 0,
    MDV = 1,
    FOOD  = 0
  )

```

# Create Base (Demographics + Dose History) Dataset
```{r}
doses_demogs <- subject_ex |> 
  left_join(subject_dm_lb, by = c("USUBJID", "COHORT")) |> 
  mutate(
    C = NA_integer_,
    LINE = row_number(),
    ID = match(USUBJID, unique(USUBJID)),
    STUDY = 305,
    LLOQ = 0.1,
    DOSEA = AMT,
    DOSEN = AMT,
    REGN = 2,
    DV = NA_integer_,
    LDV = NA_integer_
  ) |> 
  select(names(spec))

```

```{r}

write_csv(doses_demogs, 
          file = file.path(derDir, "subject_demog_dose.csv"), 
          na = ".")

digest::digest(doses_demogs) |> print()

```


# Test Simulation Run (validate dataset for use in mrgsolve and NONMEM)

```{r}
# validate test for simulation wiht dataset
sim1 <- mrgsim_df(mod1,
                  data = doses_demogs |> 
                    filter(ATFD <= 24*16) |> 
                    mutate(ETA1 = 0, ETA2 = 0, ETA3 = 0, 
                           ETA4 = 0, ETA5 = 0) |> 
                    mutate(TIME = ATFD),
                  etasrc = "data.all",
                  carry.out = names(spec))


# example subject simulation
ggplot(sim1 |> filter(ID == 47), aes(TIME, IPRED)) + 
  geom_point(col = "darkblue", shape = 1) + 
  geom_line(col = "darkblue") + 
  geom_point(aes(y = Y), col = "darkgreen", shape = 1) + 
  geom_line(aes(y = Y), col = "darkgreen") + 
  xlim(0, 120) +
  xlab("Time since first dose (hours)") + 
  ylab("ABC-XYZ Serum Concentration (ng/mL)") + 
  ggtitle("Example Subject (ID = 47, 4.8 mg) First 3 Doses")
  
```
BLUE  = individual predicted value;
GREEN = simulation of value based on known variability (residual error)
