---
title: "ABC-XYZ Premise and Dataset Creation"
author: Jonah Lyon (jonah@a2-ai.com)
date: today
date-format: long
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

The purpose of this script is to use the imaginary patient population from the previous step to simulate out some simmed "real" concentrations using the PopPK model with samples from a proposed study design.  500 replicate simulations will be created so that in the next step, we can obtain a sufficient range of estimates that can be considered for the given study design.


```{r, warning = FALSE, echo = FALSE, message = FALSE}
renv::snapshot()
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(mrgsolve)
library(yspec)

# data spec yaml file
spec <- ys_load("../data/derived/ppk.yaml")
derDir <- "../data/derived/dose_histories"
base_data <- read_csv(file.path(derDir, "subject_demog_dose.csv"), na = ".")

# UPDATE if new simulation estimation iteration
simdatDir <- "../model/nonmem/simest1"

```

# Load PopPK Model Run for ABC-XYZ
```{r}
mod1 <- mread("../model/R-mrgsolve/1001.mod", 
              start = 0, end = 120, delta = 1)

```

# Proposed Study Design
- C1D1 Samples: pre-dose, 0.5, 1, 12
- C1D15 Samples: pre-dose
- C2D1 Samples: pre-dose, 0.5, 1, 12
- C2D15 Samples: pre-dose
- C3D1, C4D1 Samples: pre-dose
## General Schedule Creating
```{r}
# create generalized PK schedule
pk_samp_sched <- 
  data.frame(
    TPT = c(-0.083, 0.5, 1, 2, 4, 12)
    ) |> 
  crossing(
    DAY = c(1, 15, 29, 43, 57, 85)
  ) |> 
  filter(ifelse(DAY == 15 | DAY > 29, TPT == -0.083, TRUE)) |> 
  arrange(DAY, TPT) |> 
  mutate(
    NTFD = (DAY-1)*24 + TPT
  )

```
## Apply schedule to each subject
```{r}
usubjid <- base_data |> distinct(USUBJID, ID)

# apply schedule for all subjects
pk_times <- usubjid |> 
  crossing(pk_samp_sched) |> 
  mutate(
    EVID = 0,
    CMT = 2,
    DV = NA_integer_,
  )

```

# NONMEM-style dataset (for simulation with mrgsolve)
## One Iteration Dataset
```{r}

pk_itr1 <- base_data |> 
  bind_rows(pk_times) |> 
  arrange(USUBJID, NTFD) |> 
  mutate(
    NTLD = ifelse(NTFD > 0 & TPT < 0, TPT + 48, TPT),
    NTLD = ifelse(EVID == 1, 0, NTLD),
    ATFD = NTFD,
    ATLD = NTLD
  ) |> 
  group_by(USUBJID) |> 
  fill(c(STUDY, ID, LLOQ, DOSEA, DOSEN, BWT, BHT, BBMI, BAGE, SEXF, RACEN,
         PTYPE, BECOG, BALB, BCRCL, BHFC, REGN, FOOD), .direction = "downup") |> 
  select(-DAY, -TPT) |> 
  ungroup() |> 
  arrange(ID, ATFD)

```

## Iterate Dataset 500 times
```{r}
n <- length(unique(pk_itr1$USUBJID))

pk_itr500 <- pk_itr1 |> 
  crossing(SIM = 1:500) |> # 500 replicate sims
  mutate(
    ID = ID + (SIM-1)*n # rename id to get new sim iteration (e.g. ETAs are resimulated)
  ) |> 
  arrange(ID, ATFD)

```

# Simulation
## Simulate and Post-Process
```{r}
pk_sim <- mrgsim_df(mod1, 
          pk_itr500 |> rename(TIME = ATFD),
          carry_out = names(pk_itr500))

pk_sim1 <- pk_sim |> 
  mutate(
    ID = ID - (SIM-1)*n, # rename ID back to compare variability allowed for any one given subject
    ATFD = TIME,
    LDV = log(DV),
    C = ifelse(ATFD < 0, "C", NA_integer_),
    DV = round(ifelse(Y>0 & EVID == 0, Y, NA_integer_), 2),
    AMT = ifelse(EVID == 0, NA_integer_, AMT),
    MDV = ifelse(!is.na(DV), 0, 1)
  ) |> 
  left_join(usubjid, by = "ID") |> 
  select(names(spec), SIM)

```

## Example Subject - ID 47, first 10 sims
```{r}

theme_set(theme_bw())
example_47 <- pk_sim1 |> 
  filter(ID == 47, SIM %in% 1:10, EVID == 0) |>
  mutate(SIM = as.factor(SIM))

ggplot(example_47, aes(ATFD, DV, group = SIM, col = SIM)) +
  geom_point(shape = 1) + 
  geom_line() + 
  ggtitle("Subject 47 Concentration-Time Profile Across 10 Replicate Simulations") + 
  xlab("Time Since First Dose (hours)") + 
  ylab("ABC-XYZ Concentration (µg/mL)")

```

```{r}

n <- 500
walk(seq(n), function(q){
  data_itr_name <- str_c("sim", q, ".csv")
  
  temp <- pk_sim1 |> filter(SIM == q) |> 
    ungroup() |> mutate(LINE = row_number())
  
  write_csv(temp, file.path(simdatDir, data_itr_name), na = ".")
})

```


